Parsing medial prefrontal cortex: A joint meta-analytic and graph-theoretic approach.

### 2018-08-27 Manuscript/Manuscript.Rmd
More methods

  Previous meta-analytic work has provided important insights on the psychological phenomena attributed to mPFC subregions. Independent coordinate-based meta-analyses have found mirroring spatial coverage in mPFC for valuation (Bartra et al., 2010) and DMN (Laird et al., 2009). Whole brain analyses have shown that while more ventral mPFC regions are increasingly associated with reward-related elements of decision making, overall mPFC strongly coactivates with DMN regions across the literature, with no clear topographical distinction between these phenomena (de la Vega et al., 2016). More specifically, focused work aimed at disentangling DMN from subjective value has determined that these systems are inseparable within mPFC (Acikalin, Gorgolewski, & Poldrack, 2017). These findings highlight the multifaceted embedding of psychological constructs in subregions of mPFC, and suggest that DMN and valuation might indeed be subserved by the same system. However, recent activation-based meta-analytic work has demonstrated considerable domain specificity in this area (Kragel et al., 2018). While informative, a key limitation in distinguishing constructs through this approach is that meta-analyses often rely on information derived from group-averaged data. Averaging has traditionally been beneficial in identifying overall trends in brain function and organization when faced with short scanning sessions per subject. However, it has been shown that mPFC is subject to considerable idiosyncratic cortical folding (Zilles, Palomero- Gallagher, & Amunts, 2013) and inter-subject functional variability (Mueller et al., 2013) compared to the rest of the brain, thus making group-averaging studies difficult to interpret.

@@

  A persistent issue in analyzing mPFC in individuals is the various factors inducing signal dropout in its more ventral segments (Logothetis, 2008). As such, there is value in examining all possible sources of covariation that compose DMN, instead of estimating topologies one seed at a time. Connectome-based analyses of resting state functional connectivity (rsFC) have been fruitful in characterizing individualized functional topologies that match task-induced activity (Gordon et al., 2017; Laumann et al., 2015; Tobyne et al., 2018). A popular approach to analyze these connectomes is to rely on graph theoretic methods, which provide an organic avenue to understand brain network dynamics (Bassett, Zurn, & Gold, 2018). From these, community detection algorithms have been successfully utilized to section brain networks into cohesive substructures (Garcia, Ashourvan, Muldoon, Vettel, & Bassett, 2018). Such communities represent clusters of network nodes that are more connected with each other than with the rest of the network (Fortunato & Hric, 2016). Among the algorithms used to estimate these communities, modularity has been widely effective in brain networks (Garcia et al., 2018). However, in attempting to refine the DMN topology at the individual level, dividing these regions into what belongs and does not belong to DMN makes bipartision through spectral partitioning a more viable option. In short, spectral partitioning performs eigenvalue decompositon on an adjusted form of a correlation matrix (Laplacian matrix). The resulting eigenvector corresponding to the second-to-last eigenvalue (i.e. the Fiedler Vector) provides values to divide the network into a positive and a negative community (Fiedler, 1975). This method can thus capture a refined group of brain regions that more cohesively share activation patterns during rest.

  In this study, we aim to subsection regions commonly attributed to both DMN and subjective value into subject-specific DMN and non-DMN partitions, quantifying the degree of topographical heterogeneity present in this network within and across individuals. This description will allow researchers to generate more precise topographic targets for future studies of decision making. We do this by capitalizing on the respective strengths of meta-analytic and subject-specific analyses of brain data. First, we define regions of interest (ROI) by identifying overlapping activation throughout the literature projected to an established brain atlas (Glasser et al., 2013). We then produce a rsFC networks of all the surface vertices within these ROIs for each individual resting state dataset from the Human Connectome Project, and estimate the extent of DMN coverage through community detection. The resulting partitions are then evaluated for consistency across methods, throughout each individual's time series, and across subjects.

@@

**Meta-analysis**
We used data from metan-alyses that gathered peak brain coordinates of activity from studies of valuation (Bartra et al., 2010) (n = 27 studies) and DMN (Laird et al., 2009) (n = 77). These represent the surviving areas post-statistical thresholding from each study. For each peak volumetric coordinate, 10mm sphere masks were generated and then projected to a standard cortical mesh (fs_average, 32,000 vertices) using FreeSurfer. The sphere sections that overlapped with the cortex were mapped to an atlas of the human cortical surface (Glasser et al., 2016). This produced a list of standardized parcels that were reported on each study.

With these lists, we computed the number of times every two parcels co-occurred across each literature. This allowed us to generate two weighted networks in which each vertex was a brain parcel, and edges correspond to the number of times each pair of them co-occurred across studies. In order to threshold the parcels that are most likely to belong to each lliterature, we bootstrapped the co-occurrences across studies and removed edges whose co-occurrences lied below the 90th percentile of this distribution. Based on this, only edges with co-occurrences higher than 8 were preserved, and the remaining unconnected vertices (brain regions) were discarded. The connectivity strength of every pairwise parcels was then calculated as the sum of their co-occurrences. In order to select regions of interest (ROI) from mPFC and other brain areas that are central to valuation and DMN literatures, we first bootstrapped the mean strength for each graph separately, and chose areas with strength above the 95th percentile of this distribution. Next, we took the intersection of these areas, such that the final ROIs were brain regions that overlapped between DMN and valuation in the literature (and thus our main target for dissociation).

**Whole Brain Parcellated connectivity**
In order to actually quantify the intrinsic connectivity of these ROIs in the brain, we acquired resting-state data from the Human Connectome Project (n = 80). Beyond the standard minimally preprocessed pipeline from HCP (), the data went through a number of additional refinements that are explained in detail elsewhere (Tobyne et al., 2017). Briefly, each dataset was bandpass filtered (DEETS HERE), and had the mean greyordinate signal regressed from every vertex time series (MGSR). Each data set contained around 1 hour of signals represented as approximately 4800 time points, and each subject contained 60,000 vertices. I parcellated each subject’s brain according to the atlas mentioned above, so that each parcel contained the mean time series from its vertices, and correlated the mean time series among all parcels (Pearson). This produced a weighted adjacency matrix for each subject, where each of the 360 parcels is a vertex, and edges the correlations among all of them. Next, I took the exponential of the correlations, Fisher transformed them (i.e. tanh) so that all weights were positive while maintaining the shape of the original correlation distribution, and retained the edges whose adjusted p-values remained significant (FDR < 0.05). Again, I quantified each parcel’s strength and betweenness, averaged them across subjects, and examined whether any of the ROIs showed uniquely high centralities. For this, I performed a permutation comparing the mean strength between the ROIs and the rest of the brain for each participant.

>>>>

  Previous meta-analytic work has provided important insights on the psychological phenomena attributed to mPFC subregions. Independent coordinate-based meta-analyses have found mirroring spatial coverage in mPFC for valuation (Bartra et al., 2010) and DMN (Laird et al., 2009). Whole brain analyses have shown that while more ventral mPFC regions are increasingly associated with reward-related elements of decision making, vmPFC strongly coactivates with DMN regions across the literature, with no clear topographical distinction between these phenomena (de la Vega et al., 2016). More specifically, focused work aimed at disentangling DMN from subjective value has determined that these systems are inseparable within mPFC (Acikalin, Gorgolewski, & Poldrack, 2017). These findings highlight the multifaceted embedding of psychological constructs in subregions of mPFC, and suggest that DMN and valuation might indeed be subserved by the same system. However, recent activation-based meta-analytic work has demonstrated considerable domain specificity in this area (Kragel et al., 2018). While informative, a key limitation in distinguishing constructs through this approach is that meta-analyses often rely on information derived from group-averaged data. Averaging has traditionally been beneficial in identifying tendencies in brain function and organization when faced with short scanning sessions per subject. However, it has been shown that mPFC is subject to considerable idiosyncratic cortical folding (Zilles, Palomero- Gallagher, & Amunts, 2013) and inter-subject functional variability (Mueller et al., 2013) compared to the rest of the brain, thus making group-averaging studies difficult to interpret.

@@

  A persistent issue in analyzing mPFC in individuals is the various factors inducing signal dropout in its more ventral segments (Logothetis, 2008). As such, there is value in examining all possible sources of covariation that compose DMN, instead of estimating topologies one seed at a time. Connectome-based analyses of resting state functional connectivity (rsFC) have been fruitful in characterizing individualized functional topologies that match task-induced activity (Gordon et al., 2017; Laumann et al., 2015; Tobyne et al., 2018). A popular approach to analyze these connectomes is to rely on graph theoretic methods, which provide an organic avenue to understand brain network dynamics (Bassett, Zurn, & Gold, 2018). From these, community detection algorithms have been successfully utilized to section brain networks into cohesive substructures (Garcia, Ashourvan, Muldoon, Vettel, & Bassett, 2018). Such communities represent clusters of network nodes that are more connected with each other than with the rest of the network (Fortunato & Hric, 2016). Among the algorithms used to estimate these communities, modularity has been widely effective in brain networks (Garcia et al., 2018). However, in attempting to refine the DMN topology at the individual level, dividing these regions into what belongs and does not belong to DMN makes dissection through spectral partitioning a more valuable option. In short, spectral partitioning performs eigenvalue decompositon on an adjusted form of a correlation matrix (Laplacian matrix). The resulting eigenvector corresponding to the second-to-last eigenvalue (i.e. the Fiedler Vector) provides values to divide the network into a positive and a negative community (Fiedler, 1975). This method can thus capture a refined group of brain regions that more cohesively share activation patterns during rest.

  In this study, we aim to subsection regions commonly attributed to both DMN and subjective value into subject-specific DMN and non-DMN partitions, quantifying the degree of topographical heterogeneity present in this network within and across individuals. This description will allow researchers to generate more precise topographic targets for future studies of decision making. We do this by capitalizing on the respective strengths of meta-analytic and subject-specific analyses of brain data. First, we define regions of interest (ROI) by identifying overlapping activation throughout the literature projected to an established brain atlas (Glasser et al., 2013). We then produce a rsFC networks of all the surface vertices within these ROIs for each individual resting state dataset from the Human Connectome Project (HCP; citation), and estimate the extent of DMN coverage through spectral partitioning. The resulting patterns are then evaluated for consistency across methods, throughout each individual's time series, and across subjects.

@@

Tentative outline:

- Meta-analysis
- fMRI data
- Network definition
- Community detection
- Evaluation

**Meta-analysis**
We used data from metan-alyses that gathered peak brain coordinates of activity from studies of valuation (Bartra et al., 2010) (n = 27 studies) and DMN (Laird et al., 2009) (n = 77). These represent the surviving areas post-statistical thresholding from each study. For each peak volumetric coordinate, 10mm sphere masks were generated and then projected to a standard cortical mesh (fs_average, 32,000 vertices) using FreeSurfer. The sphere sections that overlapped with the cortex (should I add the details of the overlap?) were mapped to an atlas of the human cortical surface (Glasser et al., 2016). This produced a list of standardized parcels that were reported on each study.

With these lists, we computed the number of times every two parcels co-occurred across each literature. This allowed us to generate two weighted networks in which each vertex was a brain parcel, and edges corresponded to the number of times each pair of them co-occurred across studies. Based on this, only edges with co-occurrences higher than 8 were preserved, and the remaining unconnected vertices (brain regions) were discarded. The connectivity strength of every pairwise parcels was then calculated as the sum of their co-occurrences. In order to threshold the parcels that are most likely to belong to each literature, we bootstrapped the connectivity strength and removed edges whose values lied below the 95th percentile of this distribution. In order to select regions of interest (ROI) from mPFC and other brain areas that are central to valuation and DMN literatures, we first bootstrapped the mean strength for each graph separately, and chose areas with strength above the 95th percentile of this distribution. Next, we took the intersection of these areas, such that the final ROIs were brain regions that overlapped between DMN and valuation in the literature (and thus our main target for dissociation).

Next, we identified areas that were solely related to either literature. We limited the search space to the parcels from Glasser et al. (2013) that were contained within the DMN parcellation defined by Yeo et al. (2011). We permuted the label of every study (i.e. DMN or valuation) 5000 times while preseving the total number of studies observed in each domain, and on each iteration stored the maximum statistic from a ROI-wise chi-squared test of proportions on the number of prmuted studies that had reported each ROI. This gave us a null distribution of chi-squared values that was used to determine which ROIs were more significantly prevalent on each literature.

**fMRI Data**
In order to quantify the intrinsic connectivity of these ROIs in the brain, we acquired resting-state data from the Human Connectome Project (n = 80, randomly sampled from the total pool) Q6 release. Beyond the standard minimally preprocessed pipeline from HCP (), the data went through a number of additional refinements that are detailed elsewhere (Tobyne et al., 2017). Briefly, each dataset was bandpass filtered (DEETS HERE), and had the mean greyordinate signal regressed from every vertex time series (MGSR). Each subject's brain was comprised of 32k vertices per hemisphere. We retained only the cortical surfaces, which resulted in 59,412 total surface vertices per subject. Resting state data was acquired over four scanning sessions that occurred over the course of two days (two sessions per day). Each resting state scan was comprised of 1200 TRs (TR = 0.720s) for a total time of 14 min 33 s per session. We used this scanning structure to estimate each subject's topology based on session, day, and overall data (4800 TRs).

**Within-ROI Connectivity**
To establish each subject's network, we selected all the vertices contained within the ROIs (n = 5,942 per subject) and computed the Pearson correlation of the time series of every pair of vertices. This produced a weighted network for each subject, where surface vertices were the nodes, and edges the correlations among all of them. Next, we took the exponential of the correlations, Fisher transformed them (i.e. tanh) so that all weights were positive while maintaining the shape of the original correlation distribution, and retained the edges whose adjusted p-values remained significant after correction for multiple tests (FDR < 0.05).
